Address Clustering Heuristics for Account-Based Blockchain Networks: An Analysis based on a Decentraland User Set

### 2023-11-07 paper.tex
Update section 4 clustering heursitics

\section{Clustering Heuristics}
In this section, I present an overview of clustering heuristics in the context of account-based blockchains and on-chain data. I will provide a detailed explanation of the mechanics and rationale behind the heuristic and assess its applicability to our case. 
In Section 5, I will implement and evaluate the heuristics I determine to be suitable for our objectives.

\subsection{Self-Authorization} 
	All three token standards require functions to allow another address to transfer tokens on behalf of the actual owner. For the ERC-20 standard, this is achieved through the \texttt{`approve`} function, accepting the spender's address and the token amount as parameters\footnote{\texttt{approve(address spender, uint256 amount)}}. The ERC-721 has two functions dedicated to approvals: One for authorizing the transfer of a specific NFT\footnote{\texttt{approve(address to, uint256 tokenId)}} and one to permit transferring all of the owner's NFTs within the collection\footnote{\texttt{setApprovalForAll(address operator, bool approved)}}. The latter function also appears in the ERC-1155 standard. The approval functionality of these tokens is primarily used in connection with smart contracts (e.g., Decentralized Exchanges or NFT Marketplaces), but can also be applied for regular EOAs. \newline
The \textit{self authorization} heuristic, proposed by \cite{FV:17}, suggests that users might authorize another address they own, possibly for risk distribution across several addresses or for testing purposes. Since all the addresses' transactions are available in our data set, I can easily employ this heuristic.

\subsection{Deposit Address Reuse}
The concept of deposit address reuse was first introduced and systematically utilized by \cite{FV:17} and subsequently adoped by \cite{wu2022tutela}. This heuristic relies on the prevalent practice of crypto exchanges generating unique deposit addresses for each user. These addresses then forward the funds to an exchange's main address. It is highly likely that multiple addresses sending funds to the same deposit address are controlled by the same entity. The key challenge of this approach lies in identifying these deposit addresses, which could either be EOAs or smart contracts. \citep{FV:17}\newline
Deposit addresses share the characteristic that they forward the received funds to a major exchange account. Two parameters are essential for deposit address detection: the discrepancy between the amount received and forwarded, as the exchange must pay for gas fees, and the time lag between receiving and forwarding funds.
However, our data set only includes the transactions or token transfers associated with the addresses. The transactions between deposit and exchange address are absent. Consequently, this clustering heuristic is not applicable and would require significantly more effort for gathering the required data.

\subsection{Airdrop Multi Participation}
Airdrops are popular token distribution mechanisms, and recipients are mostly chosen based on past protocol activity. Probably the most well known example is Uniswap's UNI airdrop, where each address that ever called the Uniswap V1 or V2 contracts until a specified cutoff date received a fixed amount of 400 UNI tokens\footnote{\url{https://blog.uniswap.org/uni}}. Especially in the early days of airdrops, these distribution mechanisms were not Sybil-resistant, leading to people creating multiple addresses in anticipation of an airdrop. However, \cite{FV:17} argues that managing the tokens on all of these addresses is impractical, which is why they are often aggregated into one address. He found around 500 same-source, fixed amount token distribution events with at least 1,000 recipients while also taking into account the block difference between the individual aidrop token transfers. He defined multi-participation when two airdrop recipients forward their tokens to a single address. Since the whole view of the airdrop distribution events is not available, I am not able to identify them within our data. One possible approach would include manual selection of known airdrops. However, this process takes a long time and is not guaranteed to cluster many addresses, since more recent airdrop events were mostly carried out with some kind of Sybil-resistance. Nevertheless, to test this approach I searched for addresses that forwarded the UNI airdrop to another address. I found eight such transfers, but none involved two addresses sending funds to a third address.

@@

To construct the network graph, I use the intra-set asset transfers described in section \ref{sec:intra-set}. Our approach differs from the methodologies adopted by \cite{Beres2020} and \cite{wu2022tutela}. On the one hand, \cite{Beres2020} collected all transactions and token transfer events of their address set but did not refine this data to only include intra-set asset transfers. On the other hand, \cite{wu2022tutela} do not operate on a specific address set but use all Ethereum transactions to build the network graph, ignoring token transfer events. After comparing these two methodologies with ours, I determined that intra-set asset transfers were the most effective for clustering a predefined set of addresses. %using node-embedding methods.

@@

I focus on the three methods that performed best in \cite{Beres2020} experiments: \textit{Diff2Vec} \citep{rozemberczki2020difftovec}, \textit{Role2Vec} \citep{ahmed2018roletovec}, and \textit{DeepWalk} \citep{perozzi2014}. Diff2Vec and DeepWalk are neighborhood-preserving methods, while Role2Vec is a structural node embedding method. Although all methods utilize random walks to capture graph information, their approaches to leveraging these random walks for constructing Euclidean feature vectors differ. A detailed explanation of each algorithm is beyond the scope of this work. The interested reader may consult the original publications. For a concise overview of Diff2Vec, please refer to \cite{wu2022tutela}. \newline

@@

It is worth noting that unlike \cite{wu2022tutela}, I decided not to include edge weights. For instance, if Alice sent Bob Ether three times and Bob sent Alice Ether two times, the edge weight would correspond to five. This choice facilitated the application of \cite{rozemberczki2020difftovec}'s library and eliminated the need to develop a custom method. Based on our network graph, I generate 128-dimensional feature vectors for each address and method. \newline

@@

For evaluation of the three node embedding algorithms, I use our ENS address pairs as ground truth. I run each node embedding method ten times and report the average rank of the target address. For comparison, I compute the mean, median, and standard deviation of the average rank for each method.

@@

\subsection{LAND Transfers}
In addition to the existing clustering heuristics, I propose our heuristic based on transfers of high-value NFTs such as LAND. I argue that, given the high value of Decentraland's land parcels, there is a strong likelihood that addresses are controlled by the same entity when LAND is transferred without any compensation. Users might perform such transfers to rearrange their LAND holdings, e.g., by upgrading to a more secure wallet. \newline
The primary challenge of this approach is to detect LAND transfers without any compensation. To accomplish this, I identify token or native asset transfers within the same transaction as a LAND transfer. Upon manually examining various implementations of LAND trades on the most popular NFT marketplace contracts, I noticed a recurring pattern: the user address appears in both the \texttt{''from''} and \texttt{''to''} fields if a trade occurs. Figure \ref{fig:LAND_example} illustrates an example of this pattern. 
In the case of pure LAND transfers, the sending and receiving addresses are the same for all transfers within a transaction. However, this approach does not consider other LAND trading schemes, including off-chain compensation or non-atomic (within multiple transactions) trade settlement.

>>>>

\section{Clustering Heuristics}
In this section, I present an overview of clustering heuristics in the context of account-based blockchains and on-chain data. I will provide a detailed explanation of the mechanics and rationale behind the heuristic and assess its applicability. 
In Section 5, I will implement and evaluate the heuristics I determine to be suitable for this work.

\subsection{Self-Authorization} 
	All three token standards require functions to allow another address to transfer tokens on behalf of the actual owner. For the ERC-20 standard, this is achieved through the \texttt{`approve`} function, accepting the spender's address and the token amount as parameters\footnote{\texttt{approve(address spender, uint256 amount)}}. The ERC-721 has two functions dedicated to approvals: One for authorizing the transfer of a specific NFT\footnote{\texttt{approve(address to, uint256 tokenId)}} and one to permit transferring all of the owner's NFTs within the collection\footnote{\texttt{setApprovalForAll(address operator, bool approved)}}. The latter function also appears in the ERC-1155 standard. The approval functionality of these tokens is primarily used in connection with smart contracts (e.g., Decentralized Exchanges or NFT Marketplaces), but can also be applied for regular EOAs. \newline
The \textit{self authorization} heuristic, proposed by \cite{FV:17}, suggests that users might authorize another address they own, possibly for risk distribution across several addresses or for testing purposes. Since all the addresses' transactions are available in our data set, this heuristic can be easily employed.

\subsection{Deposit Address Reuse}
The concept of deposit address reuse was first introduced and systematically utilized by \cite{FV:17} and subsequently adoped by \cite{wu2022tutela}. This heuristic relies on the prevalent practice of crypto exchanges generating unique deposit addresses for each user. The deposit addresses then forward the funds to an exchange's main address. It is highly likely that multiple addresses sending funds to the same deposit address are controlled by the same entity. The key challenge of this approach lies in identifying these deposit addresses, which could either be EOAs or smart contracts. \citep{FV:17}\newline
Deposit addresses share the characteristic that they forward the received funds to a major exchange account. Two parameters are essential for deposit address detection: the discrepancy between the amount received and forwarded, as the exchange must pay for gas fees, and the time lag between receiving and forwarding funds.
However, our data set only includes the transactions or token transfers associated with the addresses. The transactions between deposit and exchange address are absent. Consequently, this clustering heuristic is not applicable and would require significantly more effort for gathering the required data.

\subsection{Airdrop Multi Participation}
Airdrops are popular token distribution mechanisms, and recipients are mostly chosen based on past protocol activity. Probably the most well known example is Uniswap's UNI airdrop, where each address that ever called the Uniswap V1 or V2 contracts until a specified cutoff date received a fixed amount of 400 UNI tokens\footnote{\url{https://blog.uniswap.org/uni}}. Especially in the early days of airdrops, these distribution mechanisms were not Sybil-resistant, leading to people creating multiple addresses in anticipation of an airdrop. However, \cite{FV:17} argues that managing the tokens on all of these addresses is impractical, which is why they are often aggregated into one address. He found around 500 same-source, fixed amount token distribution events with at least 1,000 recipients while also taking into account the block difference between the individual aidrop token transfers. He defined multi-participation when two airdrop recipients forward their tokens to a single address. Since the whole view of the airdrop distribution events is not available, I am not able to identify them within our data. One possible approach would include manual selection of known airdrops. However, this process is cumbersome and not guaranteed to cluster many addresses, since more recent airdrop events were mostly carried out with some kind of Sybil-resistance. Nevertheless, to test this approach I searched for addresses that forwarded the UNI airdrop to another address. Eight such transfers were found, but none involved two addresses sending funds to a third address.

@@

To construct the network graph, I use the intra-set asset transfers described in section \ref{sec:intra-set}. My approach differs from the methodologies adopted by \cite{Beres2020} and \cite{wu2022tutela}. On the one hand, \cite{Beres2020} collected all transactions and token transfer events of their address set but did not refine this data to only include intra-set asset transfers. On the other hand, \cite{wu2022tutela} do not operate on a specific address set but use all Ethereum transactions to build the network graph, ignoring token transfer events. Comparing these two methodologies with mine, I found the intra-set asset transfers to be the most effective for clustering the address set examined in this work. %through node-embedding algorithms in my case
%a predefined set of addresses through node-embedding algorithms.

@@

I focus on the three methods that performed best in \cite{Beres2020} experiments: \textit{Diff2Vec} \citep{rozemberczki2020difftovec}, \textit{Role2Vec} \citep{ahmed2018roletovec}, and \textit{DeepWalk} \citep{perozzi2014}. Diff2Vec and DeepWalk are neighborhood-preserving methods, while Role2Vec is a structural node embedding algorithm. Although all methods utilize random walks to capture graph information, their approaches to leveraging these random walks for constructing Euclidean feature vectors differ. A detailed explanation of each algorithm is beyond the scope of this work. The interested reader may consult the original publications. For a concise overview of Diff2Vec, please refer to \cite{wu2022tutela}. \newline

@@

It is worth noting that unlike \cite{wu2022tutela}, I decided not to include edge weights. For instance, if Alice sent Bob Ether three times and Bob sent Alice Ether two times, the edge weight would correspond to five. This choice facilitated the application of \cite{rozemberczki2020difftovec}'s library and eliminated the need to develop a custom method. Based on the network graph, I generate 128-dimensional feature vectors for each address and method. \newline

@@

For evaluation of the three node embedding algorithms, I use the ENS address pairs as ground truth. I run each node embedding method ten times and report the average rank of the target address. For comparison, I compute the mean, median, and standard deviation of the average rank for each method.

@@

\subsection{LAND Transfers}
In addition to the existing clustering heuristics, I propose a heuristic based on transfers of high-value NFTs such as LAND. I argue that, given the high value of Decentraland's land parcels, there is a strong likelihood that addresses are controlled by the same entity when LAND is transferred without any compensation. Users might perform such transfers to rearrange their LAND holdings, e.g., by upgrading to a more secure wallet. \newline
The primary challenge of this approach is to detect LAND transfers without any compensation. To accomplish this, token or native asset transfers within the same transaction as a LAND transfer are identified. Upon manually examining various implementations of LAND trades on the most popular NFT marketplace contracts, I noticed a recurring pattern: the user address appears in both the \texttt{''from''} and \texttt{''to''} fields if a trade occurs. Figure \ref{fig:LAND_example} illustrates an example of this pattern. 
In the case of pure LAND transfers, the sending and receiving addresses are the same for all transfers within a transaction. However, this approach does not consider other LAND trading schemes, including off-chain compensation or non-atomic (within multiple transactions) trade settlement.
